Flowchart of Current Algorithm

### 2010-04-05 docs/flowchart/flowchart.tex
add in some equations for context

\item Compute $S^{\nph,\star\star}$.  We compute an estimate for the 
  time-centered expansion term in the velocity 
  divergence constraint (eq. [\ref{eq:utilde divergence}]).  For the  
  first time step ($n=0$), we set
\begin{equation}
S^{n+\myhalf,\star\star} = \frac{S^0 + S^1}{2},
\end{equation}
  where $S^1$ is found during initialization.  For other time steps
  $(n \ne 0)$, following \cite{Bell:2004}, we extrapolate
  to the half-time using $S$ at the previous and current
  time levels
\begin{equation}
S^{\nph,\star\star} = S^n + \frac{\dt^n}{2} \frac{S^n - S^{n-1}}{\dt^{n-1}}.
\end{equation}
  Next, compute
\begin{equation}
\overline{S^{\nph,\star\star}} = {\mathrm{\bf Avg}} \left(S^{\nph,\star\star}\right).
\end{equation}

@@

\item Compute the provisional base state velocity forcing.
  Rearrange equation (\ref{eq:w0 evolution}),
\begin{equation}
-\frac{1}{\rho_0} \frac{\partial \pi_0}{\partial r} = \frac{\partial w_0}{\partial t} + w_0 \frac{\partial w_0}{\partial r}, \label{eq:pizero}
\end{equation}
  with the following discretization:
\begin{equation}
\left ( \frac{1}{\rho_0} \frac{\partial \pi_0}{\partial r} \right )^{n,\star} = -\frac{w_0^{\nph,\star} - w_0^\nmh}{(\dt^n+\dt^{n-1})/2} - w_0^{n,\star} \left(\frac{\partial w_0}{\partial r}\right)^{n,\star},
\end{equation} 
  where $w_0^{n,\star}$ and $(\partial w_0 / \partial r)^{n,\star}$ are defined as
\begin{eqnarray}
w_0^{n,\star} &=& \frac{\dt^{n} w_0^{\nmh} + \dt^{n-1} w_0^{\nph,\star}}{\dt^n+\dt^{n-1}}, \\
\left(\frac{\partial w_0}{\partial r}\right)^{n,\star} &=& \frac{1}{\dt^n+\dt^{n-1}}\left [ \dt^{n} \left(\frac{\partial w_0 }{ \partial r}\right)^{\nmh} + \dt^{n-1} \left(\frac{\partial w_0 }{ \partial r}\right)^{\nph,\star} \right ].\nonumber \\
\end{eqnarray}
  If $n=0$, we use $\dt^{-1} = \dt^0$.
\end{enumerate}

@@

Using equation (\ref{eq:utildeupd}), we compute time-centered edge 
velocities, $\uadvonedag$, using 
$\Ub = \Ubt^n + w_0^{\nph,\star}$.  The $\dagger$ superscript refers to the 
fact that the predicted velocity field does not satisfy the divergence 
constraint.  We then construct $\uadvone$ from $\uadvonedag$
using a MAC projection, as described in detail in Appendix B of Paper III.  
We note that $\uadvone$ satisfies the discrete version of
$\overline{(\uadvone\cdot\eb_r)}=0$ as well as
\begin{eqnarray}
\nabla \cdot \left(\beta_0^n \uadvone\right) &=& \beta_0^n \left(S^{\nph,\star\star} - \overline{S^{\nph,\star\star}}\right),\\
 \beta_0^n &=& \beta_0 \left(\rho_0^n, p_0^n, \overline{\Gamma_1^n}\right),
\end{eqnarray}
where $\beta_0$ is computed as described in Appendix C of Paper III.

@@

\item Update $(\rho X_k)$ using a discretized version of equation 
(\ref{eq:species}) omitting the reaction terms, which were already 
accounted for in {\bf React State}.  The update consists of two steps:

@@

\item Compute the time-centered species edge states, $(\rho X_k)^{\nph,\star,\pred}$,
  for the conservative update of $(\rho X_k)^{(1)}$.  We use equations 
  (\ref{eq:Perturbational Density}) and (\ref{eq:Primitive Species}) to 
  predict $\rho^{'(1)} = \rho^{(1)} - \rho_0^n$ and 
  $X_k^{(1)} = (\rho  X_k)^{(1)} / \rho^{(1)}$ to time-centered edges using 
  $\Ub = \uadvone+w_0^{\nph,\star}\eb_r$, yielding $\rho^{'\nph,\star,\pred}$ 
  and $X_k^{\nph,\star,\pred}$.
  We convert the perturbational density full state density using
\begin{equation}
\rho^{\nph,\star,\pred} = \rho^{'\nph,\star,\pred} + \frac{\rho_0^n + \rho_0^{(2a),\star}}{2},
\end{equation}
  where the base state density terms are mapped to Cartesian edges.
  Then,\\
  $(\rho X_k)^{\nph,\star,\pred} = (\rho^{\nph,\star,\pred}X_k^{\nph,\star,\pred})$.

@@

\item Update the enthalpy using a discretized version of equation
(\ref{eq:enthalpy}), again omitting the reaction and heating terms
since we already accounted for
them in {\bf React State}.  This equation takes the form:
\begin{equation}
\frac{\partial (\rho h)}{\partial t}  = - \nabla \cdot (\Ub \rho h) + \psi + (\Ubt \cdot \eb_r) \frac{\partial p_0}{\partial r}.
\end{equation}
For spherical geometry, we solve the
analytically equivalent form,
\begin{equation}
\frac{\partial (\rho h)}{\partial t}  = - \nabla \cdot (\Ub \rho h) + \psi + \nabla \cdot (\Ubt p_0) - p_0 \nabla \cdot \Ubt,
\end{equation}
which experience has shown to minimize the drift from thermodynamic equilibrium.  The update consists of two steps:

@@

\item Compute the time-centered enthalpy edge state, $(\rho h)^{\nph,\star,\pred},$
  for the conservative update of $(\rho h)^{(1)}$.  We use equation 
  (\ref{eq:Perturbational Enthalpy}) to predict
  $(\rho h)' = (\rho h)^{(1)} - (\rho h)_0^n$ to time-centered edges, 
  using $\Ub = \uadvone+w_0^{\nph,\star} \eb_r$,
  yielding $(\rho h)^{'\nph,\star,\pred}$.  We convert the perturbational 
  enthalpy to a full state enthalpy using
\begin{equation}
(\rho h)^{\nph,\star,\pred} = (\rho h)^{'\nph,\star,\pred} + \frac{(\rho h)_0^n + (\rho h)_0^{n+1,\star}}{2}.
\end{equation}
  For planar geometry, we map $(\rho h)_0$ directly to Cartesian edges.
  In spherical geometry, our experience has shown that a slightly different
  approach leads to reduced discretization errors.  We first map 
  $h_0 \equiv (\rho h)_0/\rho_0$ and $\rho_0$ to Cartesian edges separately, 
  and then multiply these terms to get $(\rho h)_0$.

>>>>

\item Compute $S^{\nph,\star\star}$.  We compute an estimate for the
  time-centered expansion term in the velocity divergence constraint,
  as given in paper~III equation (19),
\begin{equation}
  S =  -\sigma  \sum_k  \xi_k \omegadot_k  + 
  \frac{1}{\rho p_\rho} \sum_k p_{X_k}  {\omegadot}_k  + \sigma \Hnuc + \sigma \Hext 
  + \frac{\sigma}{\rho}\nabla\cdot\kth\nabla T \enskip .
\label{eq:defineS} 
\end{equation}

  For the first time step ($n=0$), we set
\begin{equation}
S^{n+\myhalf,\star\star} = \frac{S^0 + S^1}{2},
\end{equation}
  where $S^1$ is found during initialization.  For other time steps
  $(n \ne 0)$, following \cite{Bell:2004}, we extrapolate
  to the half-time using $S$ at the previous and current
  time levels
\begin{equation}
S^{\nph,\star\star} = S^n + \frac{\dt^n}{2} \frac{S^n - S^{n-1}}{\dt^{n-1}}.
\end{equation}
  Next, compute
\begin{equation}
\overline{S^{\nph,\star\star}} = {\mathrm{\bf Avg}} \left(S^{\nph,\star\star}\right).
\end{equation}

@@

\item Compute the provisional base state velocity forcing, using equation (38)
  from paper~III,
\begin{equation}
-\frac{1}{\rho_0} \frac{\partial \pi_0}{\partial r} = \frac{\partial w_0}{\partial t} + w_0 \frac{\partial w_0}{\partial r}, \label{eq:pizero}
\end{equation}
  with the following discretization:
\begin{equation}
\left ( \frac{1}{\rho_0} \frac{\partial \pi_0}{\partial r} \right )^{n,\star} = -\frac{w_0^{\nph,\star} - w_0^\nmh}{(\dt^n+\dt^{n-1})/2} - w_0^{n,\star} \left(\frac{\partial w_0}{\partial r}\right)^{n,\star},
\end{equation} 
  where $w_0^{n,\star}$ and $(\partial w_0 / \partial r)^{n,\star}$ are defined as
\begin{eqnarray}
w_0^{n,\star} &=& \frac{\dt^{n} w_0^{\nmh} + \dt^{n-1} w_0^{\nph,\star}}{\dt^n+\dt^{n-1}}, \\
\left(\frac{\partial w_0}{\partial r}\right)^{n,\star} &=& \frac{1}{\dt^n+\dt^{n-1}}\left [ \dt^{n} \left(\frac{\partial w_0 }{ \partial r}\right)^{\nmh} + \dt^{n-1} \left(\frac{\partial w_0 }{ \partial r}\right)^{\nph,\star} \right ].\nonumber \\
\end{eqnarray}
  If $n=0$, we use $\dt^{-1} = \dt^0$.
\end{enumerate}

@@

The local velocity field is described by paper III equation (37),
\begin{equation}
\frac{\partial\Ubt}{\partial t} = 
- \left(\Ubt+w_0\right) \cdot \nabla \Ubt
- \left(\Ubt \cdot \eb_r\right) \frac{\partial w_0}{\partial r} \eb_r
- \frac{1}{\rho} \nabla\pi
+ \frac{1}{\rho_0} \frac{\partial \pizero}{\partial r} \eb_r
- \frac{(\rho-\rhozero)}{\rho} \; g \; \eb_r  \label{eq:utildeupd}  \enskip .
\end{equation}
From this, we compute time-centered edge 
velocities, $\uadvonedag$, using 
$\Ub = \Ubt^n + w_0^{\nph,\star}$.  The $\dagger$ superscript refers to the 
fact that the predicted velocity field does not satisfy the divergence 
constraint.  We then construct $\uadvone$ from $\uadvonedag$
using a MAC projection, as described in detail in Appendix B of Paper III.  
We note that $\uadvone$ satisfies the discrete version of
$\overline{(\uadvone\cdot\eb_r)}=0$ as well as
\begin{eqnarray}
\nabla \cdot \left(\beta_0^n \uadvone\right) &=& \beta_0^n \left(S^{\nph,\star\star} - \overline{S^{\nph,\star\star}}\right),\\
 \beta_0^n &=& \beta_0 \left(\rho_0^n, p_0^n, \overline{\Gamma_1^n}\right),
\end{eqnarray}
where $\beta_0$ is computed as described in Appendix C of Paper III.

@@

\item Update $(\rho X_k)$ using a discretized version of the species 
continuity equation,
\begin{equation}
\frac{\partial (\rho X_k)}{\partial t} + \nabla \cdot (\Ub \rho X_k) = 0 \enskip .
\label{eq:species}
\end{equation}
where we omit the reaction terms, which were already 
accounted for in {\bf React State}.  The update consists of two steps:

@@

\item Compute the time-centered species edge states, $(\rho X_k)^{\nph,\star,\pred}$,
  for the conservative update of $(\rho X_k)^{(1)}$.  We use the equations 
\begin{eqnarray}
\frac{\partial\rho'}{\partial t} &=& -\Ub\cdot\nabla\rho' - 
     \rho'\nabla\cdot\Ub - \nabla\cdot\left(\rho_0\Ubt\right),
     \label{eq:Perturbational Density}  \\
\frac{\partial X_k}{\partial t} &=& -\Ub\cdot\nabla X_k + 
     \omegadot_k. \label{eq:Primitive Species}
\end{eqnarray}
  to
  predict $\rho^{'(1)} = \rho^{(1)} - \rho_0^n$ and 
  $X_k^{(1)} = (\rho  X_k)^{(1)} / \rho^{(1)}$ to time-centered edges using 
  $\Ub = \uadvone+w_0^{\nph,\star}\eb_r$, yielding $\rho^{'\nph,\star,\pred}$ 
  and $X_k^{\nph,\star,\pred}$.
  We convert the perturbational density full state density using
\begin{equation}
\rho^{\nph,\star,\pred} = \rho^{'\nph,\star,\pred} + \frac{\rho_0^n + \rho_0^{(2a),\star}}{2},
\end{equation}
  where the base state density terms are mapped to Cartesian edges.
  Then,\\
  $(\rho X_k)^{\nph,\star,\pred} = (\rho^{\nph,\star,\pred}X_k^{\nph,\star,\pred})$.

@@

\item Update the enthalpy using a discretized version of the enthalpy
evolution equation, again omitting the reaction and heating terms
since we already accounted for
them in {\bf React State}.  This equation takes the form:
\begin{equation}
\frac{\partial (\rho h)}{\partial t}  = - \nabla \cdot (\Ub \rho h) + \psi + (\Ubt \cdot \eb_r) \frac{\partial p_0}{\partial r}.
\end{equation}
For spherical geometry, we solve the
analytically equivalent form,
\begin{equation}
\frac{\partial (\rho h)}{\partial t}  = - \nabla \cdot (\Ub \rho h) + \psi + \nabla \cdot (\Ubt p_0) - p_0 \nabla \cdot \Ubt,
\end{equation}
which experience has shown to minimize the drift from thermodynamic
equilibrium.  The update consists of two steps:

@@

\item Compute the time-centered enthalpy edge state, $(\rho h)^{\nph,\star,\pred},$
  for the conservative update of $(\rho h)^{(1)}$.  The perturbational
  enthalpy equation, neglecting reactions,
\begin{equation}
\frac{\partial(\rho h)'}{\partial t} = -\Ub\cdot\nabla(\rho h)' - 
   (\rho h)'\nabla\cdot\Ub - \nabla\cdot\left[(\rho h)_0\Ubt\right] + \Ubt\cdot\nabla p_0 
    \label{eq:Perturbational Enthalpy}.
\end{equation}
  is used to predict
  $(\rho h)' = (\rho h)^{(1)} - (\rho h)_0^n$ to time-centered edges, 
  using $\Ub = \uadvone+w_0^{\nph,\star} \eb_r$,
  yielding $(\rho h)^{'\nph,\star,\pred}$.  We convert the perturbational 
  enthalpy to a full state enthalpy using
\begin{equation}
(\rho h)^{\nph,\star,\pred} = (\rho h)^{'\nph,\star,\pred} + \frac{(\rho h)_0^n + (\rho h)_0^{n+1,\star}}{2}.
\end{equation}
  For planar geometry, we map $(\rho h)_0$ directly to Cartesian edges.
  In spherical geometry, our experience has shown that a slightly different
  approach leads to reduced discretization errors.  We first map 
  $h_0 \equiv (\rho h)_0/\rho_0$ and $\rho_0$ to Cartesian edges separately, 
  and then multiply these terms to get $(\rho h)_0$.
